### 2021-05-27 paper/agujournaltemplate.tex
Last version of paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
%
% You may copy this file and give it your
% article name, and enter your text.
%
%
% Step 1: Set the \documentclass
%
%

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

\draftfalse

%% Enter journal name below.
%% Choose from this list of Journals:
%
% JGR: Atmospheres
% JGR: Biogeosciences
% JGR: Earth Surface
% JGR: Oceans
% JGR: Planets
% JGR: Solid Earth
% JGR: Space Physics
% Global Biogeochemical Cycles
% Geophysical Research Letters
% Paleoceanography and Paleoclimatology
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resources Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
%
% ie, \journalname{Water Resources Research}

\journalname{Geochemistry, Geophysics, Geosystems}

\begin{document}

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{Depth and thickness of tectonic tremor in the northeastern Olympic Peninsula}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{A. Ducellier\affil{1}, K. C. Creager\affil{1}}

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

\affiliation{1}{University of Washington}
%(repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{Ariane Ducellier}{ducela@uw.edu}

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item We use seismic data from small-aperture arrays in the Olympic Peninsula, Washington.
\item We compute the depth of the tremor source from the stacking of horizontal-to-vertical cross-correlations of seismic data.
\item Tremor is located close to the plate boundary in a region no more than 2-3 kilometers thick.
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT and PLAIN LANGUAGE SUMMARY
%
% A good Abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.

% The Plain Language Summary should be written for a broad audience,
% including journalists and the science-interested public, that will not have 
% a background in your field.
%
% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
% see http://sharingscience.agu.org/creating-plain-language-summary/)
%
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}
Tectonic tremor has been explained as a swarm of low-frequency earthquakes, which are located on a narrow fault at the plate boundary. However, due to the lack of clear impulsive phases in the tremor signal, it is difficult to determine the depth of the tremor source and the distance of the source to the plate interface with great precision. The thickness of the tremor region is not well constrained either. The tremor may be located on a narrow fault like the low-frequency earthquakes, or distributed through a few kilometers wide low shear-wave velocity layer in the upper oceanic crust, which is thought to be a region with high pore-fluid pressure. In this study, we use the cross-correlation of the horizontal and vertical components of seismic data recorded by small-aperture arrays in the Olympic Peninsula, Washington to compute the depth and thickness of the tremor. We find that the source of the tremor is located close to the plate boundary in a region no more than 2-3 kilometers thick. The source of the tremor is thus distributed over a wider depth range than the low-frequency earthquakes. However, due to the uncertainty on the depth, it is difficult to conclude whether the source of the tremor is located in the subducting oceanic crust, in the lower continental crust just above the plate boundary, or in a layer distributed above and under the plate boundary.
\end{abstract}

%\section*{Plain Language Summary}
%[ enter your Plain Language Summary here or delete this section]

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
% \section{Introduction}
%
% The main text should start with an introduction. Except for short
% manuscripts (such as comments and replies), the text should be divided
% into sections, each with its own heading.

% Headings should be sentence fragments and do not begin with a
% lowercase letter or number. Examples of good headings are:

% \section{Materials and Methods}
% Here is text on Materials and Methods.
%
% \subsection{A descriptive heading about methods}
% More about Methods.
%
% \section{Data} (Or section title might be a descriptive heading about data)
%
% \section{Results} (Or section title might be a descriptive heading about the
% results)
%
% \section{Conclusions}

\section{Introduction}

Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases. Tectonic tremor has been explained as a swarm of small, low frequency earthquakes (LFEs) ~\cite{SHE_2007_nature}, that is small magnitude earthquakes (M $\sim$ 1) where frequency content (1-10 Hz) is lower than for ordinary small earthquakes (up to 20 Hz). The source of the LFEs is located on the plate boundary, and their focal mechanisms represent shear slip on a low-angle thrust fault dipping in the same direction as the plate interface ~\cite{IDE_2007_GRL, ROY_2014}. Due to the lack of clear impulsive phases in the tremor signal, it is difficult to determine the depth of the tremor source with the same precision. The tremor source is assumed to be located close to the plate boundary. In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations ~\cite{OBA_2002, ROG_2003}. Due to this correlation, these paired phenomena have been called Episodic Tremor and Slip (ETS). \\

The occurrence of tremor seems to be linked to low effective normal stress or high fluid pressure near the location of the source of the tremor. Indeed, ~\citeA{SHE_2006} have observed a high ratio between P-wave velocity and S-wave velocity in the subducting oceanic crust near the location of the LFEs in western Shikoku, Japan. In Cascadia, ~\citeA{AUD_2009} have computed receiver functions of teleseismic waves in Vancouver Island, and analyzed the delay times between the forward-scattered P-to-S, and back-scattered P-to-S and S-to-S conversions at two seismic reflectors identified as the top and bottom of the oceanic crust. It allowed them to compute the P-to-S velocity ratio of the oceanic crust layer and the S-wave velocity contrast at both interfaces. The very low Poisson's ratio of the layer could not be explained by the mineral composition, and they interpreted it as evidence for high pore-fluid pressure. The link between tremor and high pore fluid pressure is also supported by the influence of tidal cycles on the variations of tremor occurrence. \citeA{NAK_2008} noticed that tremor swarms often exhibit occurrences with a periodicity of about 12 or 24 h, and concluded that they are probably related to Earth tides. Their occurrence is also well correlated with time evolution of Coulomb failure stress (CFS) and CFS rate. However, they noted that tremor occurrences are advanced by a few hours relative to CFS, from which they conclude that a simple Coulomb threshold model is not sufficient to explain tremor occurrence. Instead they point out that the correlation of tremor occurrence and the CFS rate as well as the time delay between both could be reproduced by using the rate- and state-dependent friction law. ~\citeA{THO_2009} have also observed that tremor occurrence on the deep San Andreas fault are correlated with small, tidal shear stress changes. They explain it by a very weak fault zone with low effective normal stress, probably due to near-lithostatic pore pressures at the depth of the tremor source region. \\

Two scenarii could explain the generation of tectonic tremor by highly pressured fluids ~\cite{SHE_2006}. A first possibility is that tremor is generated by the movement of fluids at depth, either by hydraulic fracturing or by coupling between the rock and fluid flow. The accompanying slip could be triggered by the same fluid movement that generates the tremor or, alternatively, the fluid flow could be a response to changes in stress and strain induced by the accompanying slip. The second possibility is that tremor is generated by slow otherwise aseismic shear slip on the plate interface as slip locally accelerates owing to the effects of geometric or physical irregularities on the plate interface. Fluids would then play an auxiliary role, altering the conditions on the plate interface to enable transient slip events, without generating seismic waves directly. \\

The source of the fluids could be the dehydration of hydrous minerals within the subducting oceanic crust ~\cite{SHE_2006}. ~\citeA{FAG_2011} computed the equilibrium mineral assemblages at different P-T conditions, and compared it to the P-T path of the subducting oceanic crust in Shikoku and Cascadia. They noted that for most of the P-T path, there are no dehydration reactions and the slab remains fluid-absent, except for depths between 30 and 35 km depth for Shikoku, and depths between 30 and 40 km for Cascadia, where the mineral model predicts significant water release. These depth ranges coincide with the depth range where tremor has been observed. They concluded that abundant tremor activity requires metamorphic conditions where localized dehydration occurs during subduction, which explains why the generation of tectonic tremor is restricted to a small range of depth along the plate boundary. In subduction zones where dehydration reactions are more widely distributed, there would be a more diffuse pattern of tremor activity that would be harder to detect. \\

Large amounts of fluids could be available at the fore-arc mantle corner. First, the bending of the subducting plate at the ocean trench may introduce water in the upper oceanic mantle, resulting in extensive serpentinization ~\cite{HYN_2015}. Second, the serpentinization of the fore-arc mantle corner may decrease the vertical permeability of the boundary between the oceanic plate and the overriding continental crust while keeping a high permeability parallel to the fault. It would thus channel all the fluid updip in the subducting oceanic crust, and explain the sharp velocity contrast observed by ~\citeA{AUD_2009} on top of the oceanic crust layer. At greater depth, the large volume reduction and water release accompanying eclogitization in the subducted oceanic crust, and the large volume expansion accompanying serpentinization in the mantle wedge, could increase the permeability of the plate boundary through fracture generation. A possible cause of ETS events could thus be periodic cycles of steady pore-fluid pressure build-up from dehydration of subducted oceanic crust, fluid release from fracturing of the interface during ETS, and subsequent precipitation sealing of the plate boundary ~\cite{AUD_2009}. \\

Whereas the position of the subduction zone ETS does not seem to coincide with a specific temperature or dehydration reaction ~\cite{PEA_2009}, there seems to be a good coincidence between the location of the fore-arc mantle corner, and the location of ETS ~\cite{HYN_2015}. The generation of slow slip and tectonic tremor could then be related to the presence of quartz in the overriding continental crust. Using receiver functions of teleseismic body waves, and data from the literature, ~\citeA{AUD_2014} observed that the recurrence time of slow earthquakes increases linearly with the Vp / Vs ratio of the forearc crust. They also noticed that along a margin-perpendicular profile from northern Cascadia, the Vp / Vs ratio of the forearc, and the recurrence time of ETS events, decrease with increasing depth. Likewise, ~\citeA{HYN_2015}  pointed out that the deep fore-arc crust has a very low Poisson's ratio (less than 0.22), and that the only mineral with a very low Poisson's ratio is quartz (about 0.1), which led them to conclude that there may be a significant amount of quartz (about 10\% in volume) in the deep fore-arc crust above the fore-arc mantle. ~\citeA{AUD_2014} explained the presence of quartz in the forearc by the enrichment of forearc minerals in fluid-dissolved silica derived from the dehydration of the down-going slab. As the solubility of silica increases with temperature, fluids generated at depth and rising up the subduction channel should be rich in silica ~\cite{HYN_2015}. \\

Quartz veins have indeed been observed in exhumed subduction zones. For instance, ~\citeA{FAG_2014} have studied an exhumed shear zone representing the subduction megathrust before its incorporation into the accretionary prism. They focused their study on a 30 m high by 80 m long cliff exposure where foliation has developed as a result of shearing along the subduction thrust interface. They identified two groups of quartz veins, foliation-parallel veins, and discordant veins, that must have formed for an extended time before, during, and after foliation development. They interpret the foliation-parallel veins as having been formed by viscous shear flow, and note that the shear stain rate due to the flow may be high enough to accommodate a slow slip strain rate of $~10^{-9} s^{-1}$, for a typical subduction thrust thickness of 30 m ~\cite{ROW_2013}. They interpret the discordant veins as having been formed by brittle deformation caused by locally elevated fluid pressure. The size of the structures where brittle deformation is observed (meters to hundreds of meters) is compatible with the size of the asperity rupturing during an LFE. Tremor and slow slip may thus be a manifestation of brittle-viscous deformation in the shear zone. \\

However, several constraints remain to be explained. ~\citeA{AUD_2014} estimated that the fluid flux required for the formation of quartz veins was two orders of magnitude greater than the fluid production rates estimated from the dehydration of the slab. They hypothesized that silica-saturated fluids may originate from the complete serpentinization of the mantle near the wedge corner. They suggested that higher temperature and quartz content at depth may lead to faster dissolution - precipitation processes and more frequent slip events. Their model could also explain the global variation in recurrence time, with mafic silica-poor regions having longer ETS recurrence times than felsic silica-rich regions. \\

Moreover, the zone with high low Poisson's ratio observed by ~\citeA{HYN_2015} has a large vertical extent (about 10 kilometers). If the whole zone is associated with quartz deposition and tectonic tremor generation, we should also observe a large vertical distribution of the source of the tremor. ~\citeA{KAO_2006} have used a Source Scanning Algorithm to detect and locate tremor, and have indeed located tremor in the continental crust, with a wide depth range of over 40 km. They noted that this wide depth range could not arise from either analysis uncertainties or a systematic bias in the velocity model they used. Uncertainties on the location of the tremor have been estimated by ~\citeA{IDE_2012} at about 1.5 km in the horizontal direction and 4.5 km in the vertical direction. A follow-up study by ~\citeA{KAO_2009} gave a thickness of the tremor zone of 5-10 km. This depth range is inconsistent with the depth of the LFEs, which have been located on a thin band at or near the plate interface with a rupture mechanism that corresponds to the thrust dip angle ~\cite{IDE_2007_GRL}. In the Olympic Peninsula, LFE families have been identified and located by ~\citeA{CHE_2017_JGR}, and all LFE families were located near the plate interface. Further study is thus needed to narrow the uncertainty on the depth of the source of the tremor, and verify whether tremor could occur in a wider zone than LFEs. \\

Several methods have been developed to detect and locate tectonic tremor or LFEs using the cross-correlation of seismic signals. The main idea is to find similar waveforms in two different seismic signals, which could correspond to a single tremor or LFE recorded at two different stations, or two different tremors or LFEs with the same source location but occurring at two different times and recorded by the same station. A first method consists in comparing the envelopes of seismograms at different stations ~\cite{OBA_2002, WEC_2008}, or directly the seismograms at different stations ~\cite{RUB_2013}. For instance, ~\citeA{WEC_2008} computed the cross-correlations of envelope seismograms for a set of 20 stations in western Washington and southern Vancouver Island. Then, they performed a grid search over all possible source locations to determine which one minimizes the difference between the maximum cross-correlation and the value of the correlogram at the lag time corresponding to the S-wave travel time difference between two stations. \\

A second method is based on the assumption that repeating tremor or LFEs with sources located nearby in space will have similar waveforms ~\cite{BOS_2012, ROY_2014, SHE_2006, SHE_2007_nature}. For instance, ~\citeA{BOS_2012} looked for LFEs by computing autocorrelations of 6-second long windows for each component of 7 stations in Vancouver Island. They then classified their LFE detections into 140 families. By stacking all waveforms of a given family, they obtained an LFE template for each family. They extended their templates by adding more stations and computing cross-correlations between station data and template waveforms. They used P- and S-traveltime picks to obtain an hypocenter for each LFE template. By observing the polarizations of the P- and S-waveforms of the LFE templates, they computed focal mechanisms and obtained a mixture of strike slip and thrust mechanisms, corresponding to a compressive stress field consistent with thrust faulting parallel to the plate interface. Further study showed that the average double couple solution is generally consistent with shallow thrusting in the direction of plate motion \cite{ROY_2014}. \\

Finally, a third method uses seismograms recorded across small-aperture arrays ~\cite{GHO_2010_GRL, LAR_2009}. For instance, ~\citeA{LAR_2009} stacked seismograms over all stations of the array for each component, and for three arrays in Cascadia. They then computed the cross-correlation between the horizontal and the vertical component, and found a distinct and persistent peak at a positive lag time, corresponding to the time between P-wave arrival on the vertical channel and S-wave arrival on the horizontal channels. Using a standard layered Earth model, and horizontal slowness estimated from array analysis, they computed the depths of the tremor sources. They located the sources near or at the plate interface, with a much better depth resolution than previous methods based on seismic signal envelopes, source scanning algorithm, or small-aperture arrays. They concluded that at least some of the tremor consisted in the repetition of LFEs as was the case in Shikoku. A drawback of the method was that it could be applied only to tremor located beneath an array, and coming from only one place for an extended period of time. \\

In this study, we extend on the method used by ~\citeA{LAR_2009} using the cross-correlation between horizontal and vertical components of seismic recordings to estimate the depth of the source of the tectonic tremor, and the depth extent of the region from which tremor originates. If indeed tremor is made of swarms of LFEs, and both represent the regions of deformation during ETS events, we would expect the thickness of the tremor to be the same as the thickness of the LFEs. If not, tremor may be occurring where LFEs are harder to detect because either smaller in amplitude, or spread out over continuous space and not as clearly repeating, and thus harder to detect.

\section{Data}

The data were collected during the 2009-2010 Array of Arrays experiment. Eight small-aperture arrays were installed in the northeastern part of the Olympic Peninsula, Washington. The aperture of the arrays was about 1 km, and station spacing was a few hundred meters. The arrays were around 5 to 10 km apart from each other (Figure 1). Most of the arrays recorded data  for most of  a year, between June 2009 to September 2010, and captured the main August 2010 ETS event. The arrays also recorded the August 2011 ETS event with a slightly reduced number of stations. ~\citeA{GHO_2012} used a multibeam backprojection (MBBP) technique to detect and locate tremor. They bandpass filtered the vertical component between 5 and 9 Hz and divided the data into one-minute-long time windows. They performed beam forming in the frequency domain at each array to determine the slowness vectors, and backprojected the slownesses through a 3-D wavespeed model ~\cite{PRE_2003} to locate the source of the tremor for each time window. This produced 28902 tremor epicenters for one-minute-long time windows during June, 2009 - September, 2010 and 5600 epicenters during August - September, 2011.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 2.5cm 0cm 9.5cm},clip]{figures/arrays_location.eps}
\caption{Map showing the location of the eight arrays (black triangles) used in this study. Grey dots are the locations of the source of the tremor recorded by the arrays. Inset shows the study area with the box marking the area covered in the main map. Contour lines represent a model of the depth of the plate interface ~\cite{MCC_2006}.}
\label{pngfiguresample}
\end{figure}

\section{Method}

For each array, and every 5 km by 5 km grid cell located within 25 km, we analyze the one-minute-long time windows corresponding to all the tremor epicenters located within the grid cell. For each one-minute-long time window, we downloaded the seismic data for each seismic station of the array. Then, for each seismic station and each channel, we detrended the data, tapered the first and last 5 seconds of the data with a Hann window, removed the instrument response, bandpass filtered between 2 and 8 Hz, and resampled the data to 20 Hz. All these preprocessing operations were done with the Python package obspy. For each seismic station and each one-minute-long time window, we cross-correlated the vertical component with the East-West component and with the North-South component. Then, we stacked the cross-correlation functions over all the seismic stations of the array. We experimented with a linear stack, a nth-root stack, and a phase-weighted stack ~\cite{SCH_1997}. We then computed the envelopes of the stacked cross-correlation functions. Figure 2 shows an example of the envelopes of the cross-correlation functions for the Big Skidder array for the 110 one-minute-long time windows when tremor was detected in a 5 km by 5 km grid cell located 7 kilometers southwest of the array. We can see that for about half of the tremor windows, there is a peak in the cross-correlation at about 4.85 s. As the energy of the P-waves is expected to be higher on the vertical component, and the energy of the S-waves to be higher on the horizontal components, we assume that this peak corresponds to the time lag between the arrival of a direct P-wave and a direct S-wave. We then stacked the cross-correlation functions over all the one-minute-long time windows. Again, we experimented with a linear stack, a nth-root stack, and a phase-weighted stack. We assume that the time of the maximum absolute value of the peak of the stack is the time lag between the arrival of the direct P-wave and the arrival of the direct S-wave.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/BS_-05_-05_PWS_PWS_cluster_ccwin.eps}
\caption{Envelopes of the stacked cross-correlation functions ordered by time for the Big Skidder array for the 110 one-minute-long time windows when tremor was detected in a 5 km by 5 km grid cell located 7 kilometers southwest of the array. Top time windows in black are the time windows that fit well with the stack. Bottom time windows in blue are time windows that do not fit well with the stack. Left panel is the cross-correlation of the EW component with the vertical component, and right panel is the cross-correlation of the NS component with the vertical component. The cross-correlation functions have been cut between 2 and 8 seconds to focus on the time lags around the peaks}
\label{pngfiguresample}
\end{figure}

Only about half of the envelopes of the cross-correlation functions have a distinct peak that coincides with the peak in the stacked cross-correlation. The other cross-correlations functions show either a distinct peak at another time lag, or no clearly visible peak. This may be either because the source of the tremor during the corresponding one-minute-long time window was mislocated, or because the signal-to-noise ratio is too low. To improve the signal-to-noise ratio of the peak in the stacked cross-correlation, we divided the one-minute-long time windows into two clusters, the ones that match well the stacked cross-correlation, and the ones that do not match it well. For each one-minute-long time window, we took the envelope of the cross-correlation function between the horizontal and the vertical component, and cross-correlated it with the stack over all the time windows. From this last cross-correlation function, we then computed its maximum absolute value, its value at time 0, and the time lag at which it takes its maximum absolute value. For each one-minute-long time window, we also took the cross-correlation function between the horizontal and the vertical component and computed the ratio between the amplitude of the peak to the root mean square. We did this for both the East-West component and the North-South component. Each one-minute-long time window is thus associated to eight values of quality criteria. We then classified each one-minute-long time window into two different clusters, based on the value of these criteria, using a K-means clustering algorithm (function sklearn.cluster.KMeans from the Python library SciKitLearn). The K-means procedure is as follows: We choose the number of clusters R, then we arbitrarily choose a center for each cluster. We put each one-minute-long time window into the cluster to which it is closest (based on the values of the eight criteria). Once all one-minute-long time windows have been put in a cluster, we recompute the mean of the eight criteria for each cluster, and iterate the procedure until convergence. On average, about 35\% of the time windows fit well with the stack and are kept in the first cluster, while 65\% do not fit well with the stack and are removed. For each cluster, we then stacked the cross-correlation functions over all the one-minute-long time windows belonging to the cluster using a phase-weighted stack. We always use two clusters. We try using more clusters, but it did not improve the final stack. Figure 3 shows the stack of the envelopes of the cross-correlation for each of the clusters compared to the original stacked cross-correlation for all the one-minute-long time windows. The clustering has improved the amplitude of the peak for one of the clusters, and made the peak nearly disappear for the other cluster.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/BS_-05_-05_PWS_PWS_cluster_stackcc.eps}
\caption{Stack of the cross-correlation functions over all the 110 time windows from Figure 2. We used a phase-weighted stack for the stack over the stations and for the stack over the time windows. The black line is the stack of the cross-correlation functions over all the time windows. The red line on the left panels is the stack of the envelopes of the cross-correlation functions over the time windows in the first cluster (which contains the time windows that do not fit well with the stack), and the blue line on the right panels is the stack of the envelopes of the cross-correlation functions over the time windows in the second cluster (which contains the time windows that fit well with the stack). Top panels are the cross-correlation of the EW component with the vertical component, and bottom panels are the cross-correlation of the NS component with the vertical component. The stacked cross-correlation function has been cut between 2 and 8 seconds to focus on the time lags around the peak.}
\label{pngfiguresample}
\end{figure}

We did this analysis for grid cells located in a 50 km by 50 km area centered on each of the eight arrays. We thus have 11 * 11 * 8 = 968 values of the time lag between the arrival of the direct P-wave and the arrival of the direct S-wave. We first assumed that the epicenter is at the center of the grid and determined the depth that satisfies the observed S-minus-P time for a 1D velocity model. The velocity models are taken from the 3D velocity model from ~\citeA{MER_2020}. For each array, we looked for the closest grid point (in terms of latitude and longitude) in the 3D model by ~\citeA{MER_2020} and used the corresponding  layered model to compute the depth. We thus used 8 different 1D velocity models corresponding to the 8 arrays. This allows us to take into account the variations of the Poisson's ratio in the East-West direction.

\section{Results}

The analysis we carried out above may not lead to a reliable value of the depth of the source of the tremor for all relative positions of the array and the tremor. First, some areas are badly covered, and only a few tremor or no tremor at all were recorded. We limit our analysis to grid cells where there are at least 30 one-minute-long time windows in the best cluster. Second, the method works best for nearly vertical ray paths. A third problem is that we assumed that the location of the tremor source is fixed during the one-minute-long time window where we compute the cross-correlation of the seismic signal. However, during an ETS event, rapid tremor streaks have been observed to propagate up-dip or down-dip at velocities ranging on average between 30 and 110 km/h ~\cite{GHO_2010_G3}, which corresponds to a maximum source displacement of 0.9 km updip or downdip during the 30 seconds duration before and after the middle of the time window. The change in predicted S-minus-P time caused by changing source location by 0.9 km in the up- or down-dip direction is small for tremor along strike from the array or in the down dip direction.  However, the change in this lag time for tremor sources up-dip from the arrays exceeds one quarter of the dominant period of the tremor signal (period= 0.33 s) for tremor sources more than 18 km updip from an array. Thus, tremors beyond 18 km updip of an array may not add coherently during rapid tremor migrations. This method works best if the P-wave is cleanly recorded on the vertical component and the S-wave on the horizontals, so generally speaking the results are more robust for vertical ray paths. \\

Finally, the data from some of the arrays are very noisy, which makes it hard to see a signal emerging when stacking over the one-minute-long time windows. We chose to keep only the locations for which the ratio between the maximum value of the stack of the envelopes of the cross-correlation functions to the root mean square is higher than 5. We compute the root mean square of the cross-correlation for a time lag between 12 and 14s because we do not expect any reflected wave to arrive that late after a direct wave. The corresponding stack of the envelopes of the cross-correlation functions are shown in Figure 4 for the East-West component and the North-South component. The grey vertical line shows the theoretical time lag between the arrival of the direct P-wave and the arrival of the direct S-wave using the 1D velocity model and the plate boundary model from ~\citeA{MCC_2006}, the dashed line corresponds to the slab model from ~\citeA{PRE_2003}. There is a good agreement between the timing of the peak and the theoretical time lag for most of the locations of the source of the tremor.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={2.5cm 0.5cm 5cm 1cm},clip]{figures/BS_PWS_PWS_0.eps}
\caption{Stack of the envelopes of the cross-correlation signals for different positions of the tremor source relative to the Big Skidder array showing the cross-correlation between the East-West and the vertical components (red lines) and between the North-South and vertical components (blue lines). The theoretical S-minus-P times (grey plain line for the McCrory model, grey dashed line for the Preston model) are generally in good agreement with the peaks in cross-correlation functions. The location of the source of the tremor varies from west to east (left to right) and from south to north (bottom to top). The numbers next to each graph indicate the number of one-minute-long time windows in the best cluster. If number is colored in black, this location is kept in the following maps, if the number is colored in grey, this location is not kept in the maps. For clarity, we plotted only the part of the cross-correlation signal corresponding to timelags between 2 and 8 seconds.}
\label{pngfiguresample}
\end{figure}

We considered only the horizontal component with the maximum value of the peak. We focus on the part of the stack of the envelopes of the cross-correlation functions located between 2 seconds before and 2 seconds after the theoretical time lag if the source of the tremor was located on the plate boundary. We computed the moment centroid of this part of the stack and assumed that it was equal to the time lag between the arrivals of the direct P-wave and the direct S-wave. We then computed the corresponding depth of the source of the tremor for all the locations of the source of the tremor. We did the same analysis for the eight arrays of the experiment. The corresponding stacks of the envelopes are shown in Figures S1 to S7 in the supplementary material. For the two arrays Cat Lake and Lost Cause, there are very few source-array locations that have both enough tremor and a high ratio between the peak and the root mean square. Moreover, for the locations where we have a peak, the peak is often not very clear and stretched along the time axis. In the following, we did not use data from these two arrays. For the Port Angeles array, there is only a clear peak for a near vertical incidence of the seismic waves. We choose to keep this array in the analysis in order to have two additional data points in the westernmost region of the study area. \\

%The method used by ~\citeA{GHO_2012} to locate the tremor is based on a multibeam-backprojection (BMMP) technique, which used only vertical channels to compute the slowness vectors, and then a 1D P-wave velocity model for the backprojection. To correct for possible mislocation of the tremor, we relocated the source of the tremor using the tremor catalog from the Pacific Northwest Seismic Network (\url{https://pnsn.org/tremor}) and computed the depth of the tremor using the new relocated latitude and longitude for each of the grid points. \\

To estimate the uncertainty on the depth of the source of the tremor, we computed the width of the stack of the envelopes of the cross-correlation functions at half the maximum of the peak. We then computed the associated depth difference using the 1D velocity model. As we computed the envelopes of the cross-correlation functions before stacking, the peak is large and also the associated uncertainty on the depth. Nevertheless, about 75\% of the data points have an uncertainty under 8km (Figure 5). The uncertainties are higher for the Burnt Hill array, and are almost always lower than 8km for the five other arrays. We tried different stacking methods, both for the stacking of the correlation functions over all the seismic stations of a given array, and for the stacking of the envelopes of those stacks over the one-minute-long time windows when tremor is detected. We experimented with a linear stack, a nth-root stack, and a phase-weighted stack, and found that using a phase-weighted stack for both the stacking over the stations and the stacking over one-minute-long time windows gave the lower uncertainty on the depth of the source of the tremor.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/uncertainty_PWS_PWS.eps}
\caption{Map of the uncertainty on the depth of the source of the tremor based on the width of the stack of the envelopes of the cross-correlation functions. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

In the following, we kept only the data points for which the maximum amplitude of the stack is higher than 0.05. Figures 6, 7, 8 and 9 show respectively a map and several cross-sections of the depth of the source of the tremor. Figure 10 and 11 show a map of the distance between the source of the tremor and the plate boundary from the McCrory model and a map of the distance to the plate boundary from the Preston model, alongside with the depth of the low-frequency earthquake families observed by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. There is a good agreement between the depth of the source of the tremor, the depth of the low-frequency earthquake families, and the depth of the plate boundary. On the cross-section figures, we compare the distance of the source of the tremor to the plate boundary for different locations of the cross-section. We note that the depth of the tremor matches well the depth of the low-frequency earthquakes.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/depth_PWS_PWS.eps}
\caption{Map of the depth of the source of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. The three black lines indicate the positions of the cross-sections shown in Figures 7 to 9. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0.5cm 5cm 2cm 7.5cm},clip]{figures/section_strike_PWS_PWS_1.eps}
\caption{Cross-section showing the depth of the source of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. The black line corresponds to the plate boundary profile of ~\citeA{MCC_2006} along the southern most black line shown on Figure 6, and the dashed line to the plate boundary profile of ~\citeA{PRE_2003}. The color bar shows the distance of the tremor and the low-frequency earthquakes along the strike. Only tremor and low-frequency earthquakes less 10km away from the profile line are shown. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0.5cm 5cm 2cm 7.5cm},clip]{figures/section_strike_PWS_PWS_2.eps}
\caption{Cross-section showing the depth of the source of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. The black line corresponds to the plate boundary profile of ~\citeA{MCC_2006} along the middle black line shown on Figure 6, and the dashed line to the plate boundary profile of ~\citeA{PRE_2003}. The color bar shows the distance of the tremor and the low-frequency earthquakes along the strike. Only tremor and low-frequency earthquakes less 10km away from the profile line are shown. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0.5cm 5cm 2cm 7.5cm},clip]{figures/section_strike_PWS_PWS_3.eps}
\caption{Cross-section showing the depth of the source of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. The black line corresponds to the plate boundary profile of ~\citeA{MCC_2006} along the northern most black line shown on Figure 6, and the dashed line to the plate boundary profile of ~\citeA{PRE_2003}. The color bar shows the distance of the tremor and the low-frequency earthquakes along the strike. Only tremor and low-frequency earthquakes less 10km away from the profile line are shown. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/d_to_pb_PWS_PWS_P.eps}
\caption{Map of the distance between the plate boundary (from the McCrory model) of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. Tremor located below the plate boundary is blue, and above is red. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/d_to_pb_PWS_PWS_P.eps}
\caption{Map of the distance between the plate boundary (from the Preston model) of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}). Tremor located below the plate boundary is blue, and above is red. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

\section{Discussion}

Previous studies have shown evidence for seismic anisotropy near the slab interface ~\cite{NIK_2009} and in the overriding continental crust ~\cite{CAS_1996}. If the source of the tremor is located in or under the anisotropic layer, the time lag between the arrival of the direct P-wave and the arrival of the direct S-wave could be different whether we consider the East-West component of the North-South component. To verify whether possible anisotropy could affect our results, we computed the time difference between the timing of the maximum of the stacked of the envelopes of the cross-correlation functions between the East-West component and the vertical component, and the timing of the maximum of the stack of the envelopes of the cross-correlation functions between the North-South component and the vertical component. We also computed the associated difference in depth of the source of the tremor (see Figure 12) for all arrays and all locations of the tremor. The difference in time is generally less than 0.25 s while the difference in depth is generally lower than 2 kilometers, which is less than the uncertainty on the depth of the source of the tremor. As the time difference between the East-West component and the North-South component due to anisotropy is expected to vary depending on the relative position of the array compared to the source of the tremor, we also plotted on Figure 12 the time difference as a function of the distance from source to array, and as a function of the azimuth, as well as the corresponding depth difference. The time difference does not increase with the distance from the source to the array, and there is no seismic path orientation that gives bigger time difference. Thus, seismic anisotropy does not seem to have a significant effect on the time lags between the arrival of the direct P-wave and the arrival of the direct S-wave, and can be neglected. ~\citeA{NIK_2009} observed anisotropy in the low-velocity layer beneath station GNW, located in the eastern Olympic Peninsula, farther south than the eight seismic arrays used in this study. If the source of the tremor is located above the low-velocity layer, we do not expect seismic anisotropy to introduce a significant effect on the time lags measured with the East-West component compared with the time lags measured with the North-South component. ~\citeA{NIK_2009} locate the low-velocity layer above the plate boundary, in the lower continental crust, which is shallower than the location given by ~\citeA{BOS_2013}, who locates the low-velocity layer in the upper oceanic crust. However, even if a low-velocity layer with 5 \% anisotropy is located in the lower continental crust as indicated by ~\citeA{NIK_2009}, the resulting difference in time lags should not be higher than a few tens of seconds, and the resulting difference in depth should not be higher that a few kilometers, which is lower than the uncertainty of the depth of the source of the tremor. ~\citeA{CAS_1996} observed anisotropy in the continental crust, especially in the upper 20 kilometers. However, the anisotropy resulted in time delays of the seismic waves of no more than 0.32s for deep earthquakes (40-60 km depth) and nor more than 0.20s for shallow earthquakes (15-30 km depth). The corresponding difference in depth of the source of the tremor would thus be less than the uncertainty.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/PWS_PWS_0.eps}
\caption{Left: S-minus-P times measured from envelopes of cross-correlation functions on the East-West component minus those from the North-South component. Right: Corresponding difference in inferred tremor depths.  These are plotted as histograms (top), versus epicentral distance between tremors and arrays (middle) and tremor to array azimuth. There is no systematic signal that might be caused by anisotropy.}
\label{pngfiguresample}
\end{figure}

Several approximations have been made to compute the depth of the source of the tremor. First, we stacked over all the tremor in 5km by 5km grid cell. The depth difference corresponding to the different locations of two points inside the grid cell would be no more than 2km for the tremor located close to the array. Even if we consider the tremor up to 25km away from the array, the corresponding depth difference would be no more than 4km, which is still lower than the uncertainty. Second, the Vp / Vs ratio is not very well known. A variation of about 1\% of the Vp / Vs ratio used to compute the depth  would lead to a variation of about 1-1.5 km of the corresponding depth and the corresponding distance to the plate boundary (Figure 13).
% ~\citeA{MER_2020} presented a new seismic tomography of the Washington forearc. Based on the transections shown in the study, the Poisson’s ratio in the eastern part of the Olympic Peninsula seems to be close to the value of 0.265 that we have used to compute the depth of the tremor. However, in the western part of the Olympic Peninsula, the Poisson’s ratio could be as low as 0.24-0.25, which would correspond to a decrease of 2-3\% of the Vp / Vs ratio, and a corresponding decrease of the depth of the source of the tremor of about 3-4.5 km. This would explain why the tremor in the western part of the study area appear farther away from the plate boundary than the tremor in the easter part.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/PWS_PWS.eps}
\caption{Variations of the depth of the tremor and we decrease the Vp / Vs ratio of 1\% (left) and when we increase the Vp / Vs ratio of 1\% (right). A change in the Vp / Vs ratio corresponds to a change in the theoretical time lag between direct P-wave and direct S-wave if the source is located on the plate boundary. Thus, it changes the time interval over which we compute the centroid of the stacked envelopes, and thus the time lag we use to compute the depth.}
\label{pngfiguresample}
\end{figure}

As proposed by ~\citeA{KAO_2009}, the source of the tremor could be distributed inside a layer that is 15 or more kilometers thick. We have assumed that all the tremor within a given grid cell originate from the same depth, and averaged over all the data to get the tremor depth. However, instead of being located on the same plane near the plate boundary, the tremor may be scattered over a layer surrounding the plate boundary. To compute the thickness of this layer, for each location of the array and the source of the tremor, we computed for each one-minute-long time window for which the cross-correlation function matches well the stacked cross-correlation the time lag between the time corresponding to the maximum absolute value for the cross-correlation function and the time corresponding to the maximum absolute value for the stacked cross-correlation. We thus obtained a distribution of time lags and the corresponding distribution of depths, and we tried to estimate the scale of the interval over which the depths vary. Using the standard deviation could lead to overestimate the width of the interval, and the corresponding thickness of the tremor layer, as the standard deviation is very sensitive to outliers. Instead, we use the Qn estimator of ~\citeA{ROU_1993}, which is a more robust estimator of scale. We then used the scale of the interval over which the depth of the source of the tremor varies as an estimate of the thickness of the tremor layer (see Figure 14). The average thickness of the tremor zone is about 1.6 km, while nearly all the values of the thickness do not exceed 2.5km, which is lower than the depth extent from ~\citeA{KAO_2009}. The smallest values for the thickness are about 0.8-2 kilometers.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/Q_PWS_PWS.eps}
\caption{Map of the thickness of the tremor layer estimated from scatter of individual one-minute-long tremor windows. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

Another way of estimating the thickness of the tremor zone is to fit a plane with a linear regression from all the values of the depth and compute the residuals from the regression. Figure 15 shows a map of the residuals. Nearly 50\% of the residuals are under 2km.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/error_PWS_PWS.eps}
\caption{Map of the residuals from the linear regression fitting all the depths to a plane.}
\label{pngfiguresample}
\end{figure}

\section{Conclusion}

We developed a method to estimate the depth of the source of the tectonic tremor, and the depth extent of the region from which the tremor originates, using stacked cross-correlations of horizontal and vertical components of seismic recordings from small aperture arrays in the Olympic Peninsula, Washington. We found that the source of the tremor is located close to the plate boundary in a region no more than 2-3 kilometers thick. The source of the tremor is thus distributed over a wider depth range than the low-frequency earthquakes. However, due to the uncertainty on the depth, it is difficult to conclude whether the source of the tremor is located in the subducting oceanic crust, in the lower continental crust just above the plate boundary, or in a layer distributed above and under the plate boundary. The relative location of the source of the tremor compared to the low-velocity layer observed near the plate boundary also remains uncertain.

%%

%  Numbered lines in equations:
%  To add line numbers to lines in equations,
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%% Enter Figures and Tables near as possible to where they are first mentioned:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables;  other caption information
%  should be placed in last line of the table, using
% \multicolumn2l{$^a$ This is a table note.}
%
%----------------
% EXAMPLE FIGURES
%
% \begin{figure}
% \includegraphics{example.png}
% \caption{caption}
% \end{figure}
%
% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
% \begin{figure}
% \noindent\includegraphics[width=\textwidth]{anothersample.png}
%\caption{caption}
%\label{pngfiguresample}
%\end{figure}
%
%
% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options. This is common when trying to add a PDF figure without pdflatex.
% \begin{figure}
% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
%\caption{caption}
%\label{pdffiguresample}
%\end{figure}
%
%
% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
%

%
% ---------------
% EXAMPLE TABLE
%
% \begin{table}
% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
% \centering
% \begin{tabular}{l c}
% \hline
%  Run  & Time (min)  \\
% \hline
%   $l1$  & 260   \\
%   $l2$  & 300   \\
%   $l3$  & 340   \\
%   $h1$  & 270   \\
%   $h2$  & 250   \\
%   $h3$  & 380   \\
%   $r1$  & 370   \\
%   $r2$  & 390   \\
% \hline
% \multicolumn{2}{l}{$^{a}$Footnote text here.}
% \end{tabular}
% \end{table}

%% SIDEWAYS FIGURE and TABLE
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{figsamp}
% \caption{caption here}
% \label{newfig}
% \end{sidewaysfigure}
%
%  \begin{sidewaystable}
%  \caption{Caption here}
% \label{tab:signif_gap_clos}
%  \begin{tabular}{ccc}
% one&two&three\\
% four&five&six
%  \end{tabular}
%  \end{sidewaystable}

%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
%\begin{linenomath*}
%\begin{equation}
%y|{f} \sim g(m, \sigma),
%\end{equation}
%\end{linenomath*}

%%% End of body of article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary, Notation or Acronym section goes here:
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
%  \begin{glossary}
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \end{glossary}

%
%%%%%%%%%%%%%%
% Acronyms
%   \begin{acronyms}
%   \acro{Acronym}
%   Definition here
%   \acro{EMOS}
%   Ensemble model output statistics
%   \acro{ECMWF}
%   Centre for Medium-Range Weather Forecasts
%   \end{acronyms}

%
%%%%%%%%%%%%%%
% Notation
%   \begin{notation}
%   \notation{$a+b$} Notation Definition here
%   \notation{$e=mc^2$}
%   Equation in German-born physicist Albert Einstein's theory of special
%  relativity that showed that the increased relativistic mass ($m$) of a
%  body comes from the energy of motion of the body—that is, its kinetic
%  energy ($E$)—divided by the speed of light squared ($c^2$).
%   \end{notation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS
%
% The acknowledgments must list:
%
% >>>>	A statement that indicates to the reader where the data
% 	supporting the conclusions can be obtained (for example, in the
% 	references, tables, supporting information, and other databases).
%
% 	All funding sources related to this work from all authors
%
% 	Any real or perceived financial conflicts of interests for any
%	author
%
% 	Other affiliations for any author that may be perceived as
% 	having a conflict of interest with respect to the results of this
% 	paper.
%
%
% It is also the appropriate place to thank colleagues and other contributors.
% AGU does not normally allow dedications.

\acknowledgments
The authors would like to thank A. Ghosh for sharing his tremor catalog. This project was funded by NSF grant EAR-1358512. A.D. would like to thank the Integral Environmental Big Data Research Fund from the College of the Environment of University of Washington, which funded cloud computing resources to carry out the data analyses. The seismic recordings used for this analysis can be downloaded from the IRIS website. Most of the figures were done using GMT ~\cite{WES_1991}. The Python scripts used to analyze the data and make the figures can be found on the first author's Github account \url{https://github.com/ArianeDucellier/timelags}.

%% ------------------------------------------------------------------------ %%
%% References and Citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% \bibliography{<name of your .bib file>} don't specify the file extension
%
% don't specify bibliographystyle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{bibliography}

%Reference citation instructions and examples:
%
% Please use ONLY \cite and \citeA for reference citations.
% \cite for parenthetical references
% ...as shown in recent studies (Simpson et al., 2019)
% \citeA for in-text citations
% ...Simpson et al. (2019) have shown...
%
%
%...as shown by \citeA{jskilby}.
%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
%...has been shown \cite{jskilbye}.
%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown by \cite <e.g.,>[and others]{lewin76}.
%
% apacite uses < > for prenotes and [ ] for postnotes
% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
%

\end{document}

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

% If you are using line numbers, remember to surround
% equations with \begin{linenomath*}...\end{linenomath*}

%  To add line numbers to lines in equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}